Modelling the Retinotectal projection - can competition provide a stopping mechanism?

### 2022-02-21 paper/paper.tex
Correction to comparison in paper
Work in progress

% Use the article doc class, with an 11 pt basic font size
\documentclass[11pt, a4paper, draft]{article}
%\documentclass[11pt, a4paper]{article}

@@

\subsection*{Receptor-ligand axon-axon interactions}

Because the competition effect, $\mathbf{C}_b$ (Eq.\,\ref{e:C}), requires a
global supervisor to provide information to each branch about the location of
all other branches, we looked for a receptor-ligand interaction that could
signal between axon branches and enable them to adjust their movements away
from one another. We investigated the idea that the same receptor-ligand
interaction that leads to gradient-based movement of RGC axons on the tectum
could also act as a signal between the axons. This is plausible, because
retinal ganglion cells are known to express ephrin receptor ligands as well as
Eph receptors \citep{ephrins_on_RGCs}. We defined $N=4$ ligands, $l_i$,
expressed by the RGCs in the model according to
%
\begin{equation} \label{e:retlig}
l_i(x,y) = 1.05 + 0.26 \exp(2.3 u).
\end{equation}
%
We used exactly the same form as the receptor expression given by
Eq.\,\ref{e:retrcpt}, but with opposite gradient direction so here $u$ is
substituted by $x$ for $l_0$, $y$ for $l_1$, $(1-x)$ for $l_2$ and by $(1-y)$
for $l_3$ (see Fig.\,\ref{f:ex}C).
%
We combined chemotaxis with a multiplicative (or mass-action?) interaction,
$\mathbf{J}_b$ (scaled by a parameter $m_{\!_J}$) to define the movement
vector, $\mathbf{M}_b$, for the $\mathbf{GJ}$ model as:
%
\begin{equation} \label{e:mv3}
 \mathbf{M}_{b} = m_{\!_G} \mathbf{G}_b + m_{\!_J} \mathbf{J}_b + 0.5 \mathbf{B}_b
 %%+  m_{\!_I} \mathbf{I}_b +  m_{\!_J} \mathbf{J}_b + 0.5 \mathbf{B}_b
\end{equation}
%
$\mathbf{J}_b$ is defined as
%
\begin{equation}
% No \frac{1}{|B_b|} in this one. Really?
\mathbf{J}_b = \sum_k^{nM} \hat{\mathbf{x}}_{kb}\,Q_{\!_J}(d_{kb})
\end{equation}
%
where $Q_{\!_J}(d_{kb})$ is the signalling strength between two growth cones of
receptor-ligand interaction radius $r_{\!_J}$ a distance $d_{kb}$ from one another,
given by
%
\begin{equation}
Q_{\!_J}(d_{kb}) = \begin{cases}
     \sum_i^N F_i\,r_{i,b}\,l_{i,k}    & d_{kb} \leq 2r_{\!_J} \\
     0 & d_{kb} > 2r_{\!_J}
     \end{cases}
\end{equation}
%
The sign of $Q_{\!_J}$ (which is dependent on the values of $F_i$) determines
whether the interaction, $\mathbf{J}_b$, is repulsive or attractive. $l_{i,k}$
is the expression of ligand $i$ on branch $k$.

@@

\subsection*{Receptor-receptor axon-axon interactions}

We modelled interactions between axon growth cones due to signalling between
receptors of the same type. For these interactions, repulsion caused a
movement of branch $b$ along a unit vector, $\hat{\mathbf{x}}_{kb}$, from
branch $k$ to branch $b$ causing them to move further apart; attraction caused
the opposite movement. This interaction, $\mathbf{I}_b$, acting on branch $b$,
is given by
%
\begin{equation}
\mathbf{I}_b
= \frac{1}{|B_{b,I}|} \sum_k^{nM} \sum_i^N \hat{\mathbf{x}}_{kb}\,W_{\!_I} \qquad \mathrm{if}~r_{i,b}
/ r_{i,k} < 1.1
\end{equation}
%
the weight here is defined as a function of $r_{\!_I}$:
%
\begin{equation}
W_{\!_I}(d_{kb}) = \begin{cases}
      1 - \frac{d_{kb}}{2r_{\!_I}}   & d_{kb} \leq 2r_{\!_I} \\
     0 & d_{kb} > 2r_{\!_I}
     \end{cases}
\end{equation}
$r_{i,k}$ is the expression of receptor $i$ on branch $k$.

See Fig.\,\ref{f:GI} and Fig.\,\ref{f:GIsurg}.

>>>>

% Use the article doc class, with an 11 pt basic font size
%\documentclass[11pt, a4paper, draft]{article}
\documentclass[11pt, a4paper]{article}

@@

\subsection*{Axon-axon interactions}

Because the competition effect, $\mathbf{C}_b$ (Eq.\,\ref{e:C}), requires a
global supervisor to provide information to each branch about the location of
all other branches, we looked for a receptor-ligand interaction that could
signal between axon branches and enable them to adjust their movements away
from one another.

% First, look at I, from Simpson and Goodhill.
First, we took inspiration from Simpson and Goodhill's `I' interaction, which
is defined as:

%We modelled interactions between axon growth cones due to signalling between
%receptors of the same type. For these interactions, repulsion caused a
%movement of branch $b$ along a unit vector, $\hat{\mathbf{x}}_{kb}$, from
%branch $k$ to branch $b$ causing them to move further apart; attraction caused
%the opposite movement. This interaction, $\mathbf{I}_b$, acting on branch $b$,
%is given by
%
\begin{equation}
\mathbf{I}_b
= \frac{1}{|B_{b,I}|} \sum_k^{nM} \sum_i^N \hat{\mathbf{x}}_{kb}\,W_{\!_I} \qquad \mathrm{if}~r_{i,b}
/ r_{i,k} > 1.1
\end{equation}
%
the weight here is defined as a function of $r_{\!_I}$:
%
\begin{equation}
W_{\!_I}(d_{kb}) = \begin{cases}
      1 - \frac{d_{kb}}{2r_{\!_I}}   & d_{kb} \leq 2r_{\!_I} \\
     0 & d_{kb} > 2r_{\!_I}
     \end{cases}
\end{equation}
$r_{i,k}$ is the expression of receptor $i$ on branch $k$.

 (supp?) Fig.\,\ref{f:GI} and Fig.\,\ref{f:GIsurg}.

% Next, replicate with receptor-ligand interaction.

% OR

% Look at receptor-ligand interaction, and just point out that the

Next, we investigated the idea that the same receptor-ligand
interaction that leads to gradient-based movement of RGC axons on the tectum
could also act as a signal between the axons. This is plausible, because
retinal ganglion cells are known to express ephrin receptor ligands as well as
Eph receptors \citep{ephrins_on_RGCs}. We defined $N=4$ ligands, $l_i$,
expressed by the RGCs in the model according to
%
\begin{equation} \label{e:retlig}
l_i(x,y) = 1.05 + 0.26 \exp(2.3 u).
\end{equation}
%
We used exactly the same form as the receptor expression given by
Eq.\,\ref{e:retrcpt}, but with opposite gradient direction so here $u$ is
substituted by $x$ for $l_0$, $y$ for $l_1$, $(1-x)$ for $l_2$ and by $(1-y)$
for $l_3$ (see Fig.\,\ref{f:ex}C).
%
We combined chemotaxis with a multiplicative (or mass-action?) interaction,
$\mathbf{J}_b$ (scaled by a parameter $m_{\!_J}$) to define the movement
vector, $\mathbf{M}_b$, for the $\mathbf{GJ}$ model as:
%
\begin{equation} \label{e:mv3}
 \mathbf{M}_{b} = m_{\!_G} \mathbf{G}_b + m_{\!_J} \mathbf{J}_b + 0.5 \mathbf{B}_b
 %%+  m_{\!_I} \mathbf{I}_b +  m_{\!_J} \mathbf{J}_b + 0.5 \mathbf{B}_b
\end{equation}
%
$\mathbf{J}_b$ is defined as
%
\begin{equation}
% No \frac{1}{|B_b|} in this one. Really?
\mathbf{J}_b = \sum_k^{nM} \hat{\mathbf{x}}_{kb}\,Q_{\!_J}(d_{kb})
\end{equation}
%
where $Q_{\!_J}(d_{kb})$ is the signalling strength between two growth cones of
receptor-ligand interaction radius $r_{\!_J}$ a distance $d_{kb}$ from one another,
given by
%
\begin{equation}
Q_{\!_J}(d_{kb}) = \begin{cases}
     \sum_i^N F_i\,r_{i,b}\,l_{i,k}    & d_{kb} \leq 2r_{\!_J} \\
     0 & d_{kb} > 2r_{\!_J}
     \end{cases}
\end{equation}
%
The sign of $Q_{\!_J}$ (which is dependent on the values of $F_i$) determines
whether the interaction, $\mathbf{J}_b$, is repulsive or attractive. $l_{i,k}$
is the expression of ligand $i$ on branch $k$.

@@

%%\subsection*{Receptor-receptor axon-axon interactions}
